---
title: ""
format: 
  html:
    embed-resources: true
editor: visual
---

```{r packages}
#| echo: false
#| warning: false

library(ggdag)
library(ggplot2)
theme_set(theme_dag())

```

## Complaint DAG

This document presents a DAG (Directed Acyclic Graph), of what happens before a consumer successfully submits a complaint. The complaint process is intended to survey a wide variety of issues that consumers  might encounter in their financial lives. 

#### What is a dag. 

In these diagrams an arrow indicates the direction of causation. Each circle (node) is a potential cause. For example, you could indicate how smoking causes cancer with a DAG.

```{r}
#| echo: false
#| warning: false
#| code-fold: true

smoking <-  dagify(
  cancer ~ tar
  , tar ~ smoking
  , smoking ~ gene
  , cancer ~ gene
  , smoking ~ peer_use
  , cancer ~ pollution)

ggdag(smoking, text = T,  text_size = 2.5) 

```

{{< pagebreak >}}

## Goals

1\) Develop a common understanding of, and language for the complaint process

2\) Identify gaps in our understanding of the complaint process

3\) Identify stuff that needs to be included in the survey instrument

4\) Incorporate anything specific from the interviews we've been working on 

{{< pagebreak >}}

## The DAG

```{r dag_set}
#| code-fold: true

coords <- tibble::tribble(
  ~name,  ~x,  ~y
  , "bank_error", 1, 3
  , "events", 1, 1.6
  , "financial_situation", 1, 2.4
  , "financial_communication", 1, 1
  , "problem", 2, 2
  , "stakes", 2, 3
  , "legal_framework", 2, 3
  , "third_party", 2, 1 
  , "problem_resolution", 3, 2
  , "persistance", 3, 3
  , "intermediaries", 3, 1 
  , "intermediaries_framework", 3, 1.5
  , "search_mediators",  4, 2
  , "complaint_submission", 5, 2
)

complaining_dag <- dagify(
  complaint_submission ~ search_mediators +  intermediaries
  , search_mediators ~  problem_resolution + persistance 
  , problem ~ financial_situation + bank_error + third_party + events + financial_communication
  , problem_resolution ~ financial_situation + third_party + 
    persistance + legal_framework + intermediaries_framework
  , intermediaries_framework ~ intermediaries
  , intermediaries ~ problem
  , legal_framework ~ problem
  , financial_situation ~ events + bank_error 
  , persistance ~ legal_framework + financial_situation
  , events ~ third_party
  , labels = c(
    "complaint_submission" = "Submission of complaint"
    , "financial_situation" = "Financial Situation"
    , "problem" = "Problem to solve"
    , "third_party" = "Third Party Causes"
    , "events" = "Unfortunate Events"
    , "problem_resolution" = "FI Resolution"
    , "persistance" = "Persistance/self-efficacy"
    , "bank_error" = "FI Error"
    , "search_mediators" = "Search process"
    , "legal_framework" = "Legal Framework"
    , "stakes" = "Consumer Stakes"
    , "intermediaries" = "Intermediaries"
    , "intermediaries_framework" = "Intermediaries Framework"
    , "financial_communication" = "Financial Communication"
    )
  , coords = coords
  , latent = c("persistance")
)

```

```{r}
#| code-fold: true
#| column: screen
#| out-width: 80%
#| fig-format: svg
#| fig-align: center

ggdag(complaining_dag, text = FALSE, use_labels = "label", text_size = 3) 

```

{{< pagebreak >}}

## Define the nodes

FI Error

:   The financial institution causes a problem for the consumer.

    **Example:** Mortgage servicing transfer

Financial Situation

:   The consumers financial health. Can be a direct cause of problems through financial mismanagement.

    **DATA:** FICO score, or financial well-being score

Unfortunate events

:   Bad stuff that happens to people but isn't directly financial. Divorce, death, job loss.

Third party causes

:   Merchant issues, criminal fraud, other financial institutions. A common example is a furnisher can cause you to try to solve a problem with a Credit Reporting Company.

Financial Communication

:   Consumer doesn't have access to or understand information about their account. Company is unresponsive, obtuse or stonewalling.

Legal Framework

:   Your problem has a relationship to legal procedures and rights. Even if that relationship is that you don't have any rights in a given situation.

Problem to solve

:   The consumer is trying to fix a specific thing. Lots of different examples--fees, confusion, data stuff, problems making payments, problems with specific transactions. 

FI resolution

:   If the FI resolves the consumer's problem in a satisfactory way then probably you won't get a complaint.

Persistence/self-efficacy

:   This is intended to capture factors that may cause someone to pursue a problem more or less intensely. This includes monetary stakes as well as psychological factors, beliefs, trusting government. Lots of stuff. This is obviously oversimplified. Arguably it should be broken in to two pre and post company interaction.

Intermediaries

:   Consumer doesn't go directly to the CFPB but goes to a third party that will contact the company or the Bureau on the Consumers Behalf.

Search process

:   This is captures difficulties in finding the correct person to complaint to. For the sake of this simple exercise I'm including the stuff that happens once the consumer lands on our submission form. If the consumer abandons their complaint, it's summarized here, although conceptually it might make more sense to break it out.
