---
title: "complaints DAG"
format: html
editor: visual
---

## Complaint Process DAG

**Inciting incident**

-   financial institution errors
-   life events
-   financial mismanagement
-   third parties (fraud, financial abuse, merchant stuff)

**Company behavior**

-   customer service
-   resolution direction (relief or no relief)
-   appropriateness of resolution (consumer is defrauded of a large sum, but only gets overdraft refunded)
-   quality of communication
-   non-customer service delays
-   company behavior is impacted by regulations

**Consumer motivations to complaint**

-   monetary stakes
-   other stakes (stay in home)
-   voice (want to change the companies policy)
    -   revenge - want to damage the company
    -   justice - want the company to do the right thing
-   control information collection/sharing/use
    -   consumer wants to control data and use of data
-   consumer understanding of legal rights
-   past experience complaining

**Consumer search**

-   internet search
-   social search
-   IRL social interaction
-   CFPB marketing
-   third parties and third party marketing (Credit repair, lawyer, etc)

Consumer then weighs one or more options and determines which one fits with motivation.

**Barriers to submission**

-   language
    -   comprehension
    -   fluency
-   time burden
-   context (landing pages, explanation)

Can we write down a dag for this?

```{r}


library(ggdag)
library(ggplot2)
theme_set(theme_dag())

complaining_dag <- dagify(
  complaint_submission ~ problem_resolution
  , problem_resolution ~ problem
  , problem ~ financial_situation
  , problem ~ bank_error
  , problem ~ third_party
  , problem ~ events 
  , financial_situation ~ ~ bank_error 
  , problem_resolution ~ financial_situation + third_party + persistance
  , financial_situation ~ events
  , persistance ~ events
  , persistance ~ stakes 
  , stakes ~ financial_situation
  , stakes ~ events
  , labels = c(
    "complaint_submission" = "Submission of complaint"
    , "financial_situation" = "Financial Situation"
    , "problem" = "Problem to complaint about"
    , "third_party" = "Third Party Actions"
    , "events" = "Unfortunate Events"
    , "problem_resolution" = "FI Resolution"
    , "persistance" = "Persistance/self-efficacy"
    , "stakes" = "Stakes"
    , "bank_error" = "Bank Error"
    )
)

ggdag(complaining_dag, text = FALSE, use_labels = "label")

```

Build can we do some simulated data too

```{r}

# this is right? or is it 
rbern <- \(n,p)(runif(n) > p)

complaints <- rbern(1000, .5)




```
